---
title: "Napa-sulfur-isotopes-analysis"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Code for analyzing sulfur isotope data comparing vineyard agriculture and non-agricultural areas of the Napa River Watershed, CA, USA

## Description
This repository includes data visualization and analysis code used for the manuscript:
Hermes, A.L., Dawson, T.E., and Hinckley, E.-L.S. (submitted March 2021) Sulfur isotopes fingerprint new aspects of sulfur cycling in human-dominated systems.

## Setup
```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)
library(dplyr)
library(tidyr)
library(readxl)
library(reshape)
library(data.table)

require("knitr")

setwd("D:/Google Drive/Research/HermesA_Research/MasterDataSheets/Napa/Packages/MAP")

```

## Import and clean data
A number of input datasets are required that are either provided in this repository or can be downloaded from public domain sources.

### watershed-wide sulfur isotopes (this study)
These data were generated for this manuscript.

Download "NapaSulfurIsotopes_v1.csv" from:
https://doi.org/10.6073/pasta/8b81b39d87d5f70325420294ffc83ddf

```{r import-watershed-S-isotopes}

# import data table
all <- read.csv("NapaSulfurIsotopes_v1.csv")

# check it out!
summary(all)

# update class types
all$site_name <- as.factor(all$site_name)
all$sample_type <- as.factor(all$sample_type)
all$prcnt_vineyard <- as.numeric(all$prcnt_vineyard)
all$prcnt_native <- as.numeric(all$prcnt_native)
all$prcnt_other <- as.numeric(all$prcnt_other)
all$sulfate_mgSperL <- as.numeric(all$sulfate_mgSperL)
all$d34S <- as.numeric(all$d34S) 
all$WY <- as.factor(all$WY)
all$sample_collection_date <- as.Date(all$sample_collection_date)


summary(all)

```

Remove outlier
```{r minus-outlier}
# remove an irrigation water sample collected during a spurious event and d34S NA vals
allno <- filter(all, d34S < 22)

```

Make a data table without the S-fungicide data
```{r remove-S-fungicide}
# S fungicide numbers will plot separately - remove for sulfate-d34s map figures
allnoF <- filter(allno, sample_type != "sulfur fungicide")

```

### lab soil leachate sulfur isotopes (prior study)
These data are from the following citation:
Hermes, A.L., Ebel, B.A., Murphy, S.F., and E.-L.S. Hinckley (2021). Fates and fingerprints of sulfur and carbon following wildfire in economically important croplands of California, U.S. Science of the Total Environment, 750(1): 142179.
https://doi.org/10.1016/j.scitotenv.2020.142179

Data are available through the Environmental Data Initiative.
Download the "napa_fire_leachingexpisotopes.csv" table from:
https://doi.org/10.6073/pasta/4473046a81e564b5bc0ccedcc2def131

```{r import-lab-soil-leachate-S-isotopes}

# import data table
leach <- read.csv("napa_fire_leachingexpisotopes.csv")

# check it out!
summary(leach)

# rename columns
colnames(leach)[6] <- "d34S"
colnames(leach)[7] <- "sulfate_mgSperL"

# only keep grassland and vineyard (not the burned samples)
leachfilt <- filter(leach, burn_unburn == "unburned")

# rename "grassland" to "native"
leachfilt$site_type[leachfilt$site_type == "grassland"] <- "native"

# create a % vineyard column
leachfilt$prcnt_vineyard <- as.numeric(as.factor(leachfilt$site_type))
leachfilt$prcnt_vineyard[leachfilt$prcnt_vineyard == 1] <- 0
leachfilt$prcnt_vineyard[leachfilt$prcnt_vineyard == 2] <- 100

```

### intensive source sulfur isotopes (prior study)
These data are from the following citation:
Hinckley, E.-L.S., Kendall, C., and Loague, K. (2008) Not all water becomes wine: Sulfur inputs as an opportune tracer of hydrochemcial losses from vineyards. Water Resources Research, 44(7): W00401.
https://doi.org/10.1029/2007WR006672

Download the "hinckley-etal-2008-WRR.csv" table from this Github repository.

```{r import-Hinckley-data}

# import data table
hinckley <- read.csv("hinckley-etal-2008-WRR.csv")

# check it out!
summary(hinckley)

# convert sample_type to a factor
hinckley$sample_type <- as.factor(hinckley$sample_type)

# convert sulfate as sulfate to sulfate as sulfur
hinckley$sulfate_mgSperL <- hinckley$sulfate_ppm * 32.065/96.06

```

## Plot data

### Manuscript Figure 2a
This figure plots sulfate concentration vs. sulfur isotope values.
Sample type is designated by point shape.
Land use/land cover category is designated by color.

```{r plot-Fig2a}

colorLULC_discrete <- ggplot() +
  # layer in all samples except the Napa River
  geom_point(data = filter(allnoF, lulc_abbrev != "napa river"),
                     aes(sulfate_mgSperL,
                         d34S,
                         shape = sample_type,
                         color = lulc_abbrev),
             stroke = 1.1, size = 4) +
  # then add in Napa River points so that they show up over the rest
  geom_point(data = filter(allnoF, lulc_abbrev == "napa river"),
             aes(sulfate_mgSperL,
                 d34S,
                 shape = sample_type,
                 color = lulc_abbrev),
             stroke = 1.1,
             size = 4) +
  scale_shape_manual(values = c(5, 8, 4, 1, 17)) +
  scale_color_manual(values = c("darkorange1","black","springgreen4","violetred")) +
  labs(x = expression(paste("SO"[4]^2^-{},"(mg S L"^-1,")")),
       y = expression(paste(delta^{34}, "S-SO"[4]^2^-{}, "(\u2030)")),
       shape = "Sample Type:",
       color = "LULC:") +
  scale_x_continuous(limits = c(0, 80), breaks = seq(0, 80, 10), expand = expansion(mult = c(0.02,0))) + 
  scale_y_continuous(limits = c(-10, 20), breaks = seq(-10, 20, 5), expand = expansion(mult = c(0.02,0.02))) +
  theme_bw() +
  theme(axis.title.x = element_text(size = 20),
        axis.text.x  = element_text(size = 20,
                                    color = "black"),
        axis.title.y = element_text(size = 20),
        axis.text.y = element_text(size = 20,
                                   color = "black"),
        plot.title = element_text(size = 20),
        legend.title = element_text(size = 20),
        legend.position = "right",
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        rect = element_rect(fill = "transparent"),
        panel.background = element_rect(fill = "transparent", color = NA),
        plot.background = element_rect(fill = "transparent", color = NA)) +
  theme(legend.text = element_text(size = 16)) +
  theme(legend.text.align = 0)

# take a look!
colorLULC_discrete

```

```{r save-Fig2a}
ggsave(colorLULC_discrete, filename = "sulfate-d34S_colorLULC.png",
       width = 8,
       height = 4,
       bg = "transparent")
```

### Manuscript Figure 2b
This figure adds in laboratory soil leachate data and plots a subset of samples to depict sulfur sources and processing.
As in Fig 2a, sample type is designated by point shape.
Land use/land cover category is designated by color.

NOTE: This figure was annotated with arrows in the printed manuscript.
```{r plot-Fig2b}

# First determine x and y ranges for precipitation and irrigation water

# From Hinckley et al. (2008)
hinckley %>% group_by(sample_type) %>% summarise(minSO4 = min(sulfate_mgSperL, na.rm = TRUE),
                                                 maxSO4 = max(sulfate_mgSperL, na.rm = TRUE),
                                                 mind34S = min(d34S, na.rm = TRUE),
                                                 maxd34S = max(d34S, na.rm = TRUE))

# From this study
allno %>% group_by(sample_type) %>% summarise(minSO4 = min(sulfate_mgSperL, na.rm = TRUE),
                                                 maxSO4 = max(sulfate_mgSperL, na.rm = TRUE),
                                                 mind34S = min(d34S, na.rm = TRUE),
                                                 maxd34S = max(d34S, na.rm = TRUE))

# And the median and range of S fungicide values
all %>% group_by(sample_type) %>% summarise(mind34S = min(d34S, na.rm = TRUE),
                                            maxd34S = max(d34S, na.rm = TRUE),
                                            meand34S = mean(d34S, na.rm = TRUE),
                                            mediand34S = median(d34S, na.rm = TRUE))

# Then make the plot
zoomin <- ggplot() +
  # First plot a shaded box that encompasses all precip/irrigation sulfur "source" values
  annotate("rect",
           xmin = 0.16,
           xmax = 3.24,
           ymin = 2.44,
           ymax = 7.47,
           alpha = 0.5,
           fill = "gray") +
  # Then plot the S fungicide values
  # median = 3.09, min = 0.95, max = 4.97
  annotate("pointrange",
           x = 0,
           y = 3.09,
           ymin = 0.95,
           ymax = 4.97,
           color = "saddlebrown",
           size = 1) +
  # Then add laboratory soil leachate as a laboratory-based "endmember"
  geom_point(data = leachfilt,
             aes(sulfate_mgSperL,
                 d34S,
                 color = site_type),
             shape = 3,
             size = 3) +
  # And finally a subset of data that shows mixing in between
  geom_point(data = filter(allnoF,site_name %in% c("Valley vineyard 1",
                                                        "Valley vineyard 2",
                                                        "Moore Creek", 
                                                        "To Kalon Creek",
                                                        "grassland",
                                                        "Huichica Creek")),
                     aes(sulfate_mgSperL,
                         d34S,
                         shape = sample_type,
                         color = lulc_abbrev),
             stroke = 1.1,
             size = 4) + 
  scale_shape_manual(values = c(5, 4, 17)) +
  scale_color_manual(values = c("darkorange1","springgreen4","violetred")) +
  labs(x = expression(paste("SO"[4]^2^-{},"(mg S L"^-1,")")),
       y = expression(paste(delta^{34}, "S-SO"[4]^2^-{}, "(\u2030)")),
       shape = "Sample Type:",
       color = "LULC:") +
  scale_x_continuous(limits = c(0, 80), breaks = seq(0, 80, 10), expand = expansion(mult = c(0.02,0))) + 
  scale_y_continuous(limits = c(-5, 20), breaks = seq(-5, 20, 5), expand = expansion(mult = c(0.02,0.02))) +
  theme_bw() +
  theme(axis.title.x = element_text(size = 20),
        axis.text.x  = element_text(size = 20,
                                    color = "black"),
        axis.title.y = element_text(size = 20),
        axis.text.y = element_text(size = 20,
                                   color = "black"),
        plot.title = element_text(size = 20),
        legend.title = element_text(size = 20),
        legend.position = "right",
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        rect = element_rect(fill = "transparent"),
        panel.background = element_rect(fill = "transparent", color = NA),
        plot.background = element_rect(fill = "transparent", color = NA)) +
  theme(legend.text = element_text(size = 16)) +
  theme(legend.text.align = 0)

# take a look!
zoomin

```

```{r save-Fig2b}

ggsave(zoomin, filename = "sulfate-d34S_subset.png",
       width = 8,
       height = 4,
       bg = "transparent")

```

### Manuscript Supplementary Figure 1
This figure depicts two points:
1) that sulfur isotopes do not appear to be an artifact of soil lysimeter sampling methods, and
2) that the pattern of sulfur isotope/concentration values is robust across vineyards and over time.

It combines new data from this study with data from Hinckley et al. (2008).

```{r plot-FigS1}

#### Merge datasets ####

# Subset data from this study
newsoilwater <- filter(allnoF, sample_type == "soil water")
newsoilwater <- newsoilwater[,c(10,15,16)]
# names(newsoilwater)[2] <- "sulfate_asS_mgL"

# Subset data from Hinckley et al. (2008)
oldsoilwater <- hinckley[,c(4,6,7)]

# Combine
allsoilwater <- full_join(newsoilwater,oldsoilwater)


#### Plot ####

allsoilwaterplot <- ggplot() +
  geom_point(data = allsoilwater,
                     aes(sulfate_mgSperL,
                         d34S,
                         shape = sample_type,
                         color = sample_type),
             stroke = 1.1, size = 4) +
  # add a logarithmic curve
  stat_smooth(data = allsoilwater,
                     aes(sulfate_mgSperL,
                         d34S),
              method = "lm",
              formula = y ~ log(x),
              se = FALSE,
              color = "black",
              linetype = "dashed") +
  scale_shape_manual(name = "Sample Type",
                     values = c(4, 2, 20, 21, 8)) +
  scale_color_manual(name = "Sample Type",
                     values = c("violetred","royalblue2","darkgrey","black","goldenrod1")) +
  labs(x = expression(paste("SO"[4]^2^-{},"(mg S L"^-1,")")),
       y = expression(paste(delta^{34}, "S-SO"[4]^2^-{}, "(\u2030)"))) +
  #scale_x_continuous(limits = c(0, 80), breaks = seq(0, 80, 10)) + 
  #scale_y_continuous(limits = c(0, 20), breaks = seq(0, 20, 5)) +
  theme_bw() +
  theme(axis.title.x = element_text(size = 20),
        axis.text.x  = element_text(size = 20,
                                    color = "black"),
        axis.title.y = element_text(size = 20),
        axis.text.y = element_text(size = 20,
                                   color = "black"),
        plot.title = element_text(size = 20),
        legend.title = element_text(size = 20),
        legend.text = element_text(size = 20),
        legend.position = "right",
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        rect = element_rect(fill = "transparent"),
        panel.background = element_rect(fill = "transparent", color = NA),
        plot.background = element_rect(fill = "transparent", color = NA)) +
  theme(legend.text.align = 0)

# take a look!
allsoilwaterplot

```

```{r save-FigS1}

ggsave(allsoilwaterplot, filename = "sulfate-d34S_all-methods.png",
       width = 12,
       height = 6,
       bg = "transparent")

```

# Statistics

## Calculate summary statistics

```{r summary-stats}

# Calculate summary statistics by LULC type
summarystats <- allnoF %>% group_by(lulc_abbrev) %>% summarise(meanSO4 = mean(sulfate_mgSperL),
                                                               stdSO4 = sd(sulfate_mgSperL),
                                                               medianSO4 = median(sulfate_mgSperL),
                                                               iqrSO4 = IQR(sulfate_mgSperL),
                                                               minSO4 = min(sulfate_mgSperL),
                                                               maxSO4 = max(sulfate_mgSperL),
                                                               meand34S = mean(d34S),
                                                               stdd34S = sd(d34S),
                                                               mediand34S = median(d34S),
                                                               iqrd34S = IQR(d34S),
                                                               mind34S = min(d34S),
                                                               maxd34S = max(d34S),
                                                               N = n())

summarystats$rangeSO4 <- summarystats$maxSO4 - summarystats$minSO4
summarystats$ranged34S <- abs(summarystats$maxd34S - summarystats$mind34S)

summarystats


# Calculate summary stats for Hinckley et al. (2008) data
hinckley %>% group_by(sample_type) %>% summarise(meanSO4 = mean(sulfate_mgSperL),
                                                 sdSO4 = sd(sulfate_mgSperL),
                                                 medianSO4 = median(sulfate_mgSperL),
                                                 iqrSO4 = IQR(sulfate_mgSperL),
                                                 meand34S = mean(d34S),
                                                 sdd34S = sd(d34S),
                                                 mediand34S = median(d34S),
                                                 iqrd34S = IQR(d34S),
                                                 N = n())

# Calculate summary stats for S fungicide samples
allno %>% filter(sample_type == "sulfur fungicide") %>% summarise(mean = mean(d34S),
                                                             median = median(d34S),
                                                             iqr = IQR(d34S),
                                                             min = min(d34S),
                                                             max = max(d34S),
                                                             N = n())

# Calculate summary stats for Hermes et al. (2021) soil leachate
leachfilt %>% group_by(site_type) %>% summarise(median = median(d34S, na.rm = TRUE),
                                           iqr = IQR(d34S, na.rm = TRUE),
                                           N = n())
```

## Analyze data
The ideal test for this dataset is a manova, which would test for differences in the multivariate d34S and sulfate concentrations across LULC groups. 

However, the data violate manova assumptions. First, a number of data points come from the same location at multiple points in time (non-independent), and second, the data are not normally distributed. 

Thus, we use non-parametric statistical analyses. We first use a Kruskal Wallis rank sum test with adjusted p-values followed by Dunn's tests to determine differences across groups.

For all statistical tests, we considered a p-value of < 0.05 significant.

```{r non-parametric-tests}

# First make lulc factor
allnoF$lulc_abbrev <- as.factor(allnoF$lulc_abbrev)

levels(allnoF$lulc_abbrev)
# levels are "mixed", "napa river", "native", and "vineyard"


#### Are differences in d34S across groups? ####
kruskal.test(d34S ~ lulc_abbrev,
             data = allnoF)

# Kruskal-Wallis rank sum test
# 
# data:  d34S by lulc_abbrev
# Kruskal-Wallis chi-squared = 55.768, df = 3, p-value = 4.708e-12

# YES, there are differences across groups since the p-value is << 0.05

# Follow with a post-hoc test with adjusted p-value to find what the differences are

# The Dunn test
dunnTest(d34S ~ lulc_abbrev,
         data = allnoF,
         method = "holm")
#              Comparison          Z      P.unadj        P.adj
# 1    mixed - napa river -3.1928356 1.408831e-03 4.226494e-03  Diff
# 2        mixed - native  1.0262457 3.047758e-01 6.095515e-01  Not Diff
# 3   napa river - native  4.4809773 7.430201e-06 3.715100e-05  Diff
# 4      mixed - vineyard -4.3628198 1.283967e-05 5.135867e-05  Diff
# 5 napa river - vineyard -0.1112654 9.114059e-01 9.114059e-01  Not Diff
# 6     native - vineyard -6.7067004 1.990743e-11 1.194446e-10  Diff


#### Are differences in [SO4] across groups? ####
kruskal.test(sulfate_mgSperL ~ lulc_abbrev,
             data = allnoF)

# Kruskal-Wallis rank sum test
# 
# data:  final_sulfate_mgSperL by lulc_abbrev
# Kruskal-Wallis chi-squared = 21.226, df = 3, p-value = 9.448e-05

# YES, there are differences across groups since the p-value is << 0.05

# Follow with a post-hoc test with adjusted p-value to find what the differences are

# The Dunn test
dunnTest(sulfate_mgSperL ~ lulc_abbrev,
         data = allnoF,
         method = "holm")
#             Comparison          Z      P.unadj        P.adj
# 1    mixed - napa river  0.8659865 3.864975e-01 1.0000000000  Not Diff
# 2        mixed - native  3.5412571 3.982254e-04 0.0019911270  Diff
# 3   napa river - native  2.2766814 2.280526e-02 0.0912210252  Not Diff
# 4      mixed - vineyard  0.3669021 7.136921e-01 0.7136920766  Not Diff
# 5 napa river - vineyard -0.7044557 4.811490e-01 0.9622980079  Not Diff
# 6     native - vineyard -4.2878759 1.803898e-05 0.0001082339  Diff

```
